---
title: "Fantasy Bachelorette 2022"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
editor_options: 
  chunk_output_type: console
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(ggtext)
library(glue)
library(shiny)
```

```{r global, include = FALSE}

# load data
  # sheet 1: includes each team, their picks, and their points
  df.teams <- readxl::read_excel("Fantasy Bachelorette, 2022_wk8.xlsx")
  
  # sheet 2: includes the bachelorette contestants and their points per week
  df.scores <- readxl::read_excel("Fantasy Bachelorette, 2022_wk8.xlsx", sheet = 2)
  
  # old, as of 2022-09-15
  # # manually  track who was eliminated and when
  # # "eliminated" column is the week they were eliminated (as an integer),
  #   # or 999 if not yet eliminated
  # df.elim <- data.frame(pick = df.scores$Participant,
  #                     eliminated = 999)
  # elim.wk1 <- c("Roby", "Joey", "Justin Y.")
  # elim.wk2 <- c("Brandan", "Chris", "Colin", "John", " Jordan V.", "Justin B.",
  #               "Matt", "Ryan")
  # # 2 column df, bachelorette contestant name, week they were eliminated
  # df.elim <- df.elim %>%
  #   mutate(eliminated = case_when(
  #     pick %in% elim.wk1 ~ 1,
  #     pick %in% elim.wk2 ~ 2,
  #     TRUE ~ eliminated
  #   ))
  
  # automatically detect week each participant was eliminated
    # pre-week 8, so all week 8 values NA, so if first NA is week 8, set to 999
    # to indicate not yet eliminated
  df.elim <- df.scores %>%
    pivot_longer(-Participant) %>%
    group_by(Participant) %>%
    summarise(eliminated = detect_index(value, is.na) - 1) %>% 
    mutate(eliminated = ifelse(eliminated == 7, 999, eliminated)) %>%
    rename(pick = Participant)
  
  # specify exceptions
  df.elim <- df.elim %>%
    mutate(eliminated = ifelse(pick == "Aven", 999, eliminated))

  
# clean and prep team data
  # get long df of each person's team
  # 2 cols: person, pick
  # 6 rows per person (1 for each of their 6 picks)
  df <- df.teams %>%
    select(!matches("\\.\\.\\d*")) %>% # get rid of columns in between names
    slice_head(n = 6) %>% # keep the 6 picks, delete the Total row
    pivot_longer(cols = everything(), names_to = "person", values_to = "pick") %>%
    arrange(person)

  # merge in weekly (and total) scores per bachelorette contestant (pick)
  df2 <- df %>% # naming variables is hard
    left_join(df.scores, by = c("pick" = "Participant"))

  # create df of total score per team, add rank for each person
  team.scores <- df2 %>%
    group_by(person) %>%
    summarise(score = sum(Total)) %>%
    arrange(-score) %>%
    mutate(rank = rank(-score, ties.method = "min"))
  
  # sort contestants from high score to low score (total score)
  # "bguys" is the bachelorette contestants, for short
  bguys.order <- fct_reorder(df.scores$Participant, df.scores$Total)

  # sort team scores from high to low total score
  team.order <- fct_reorder(team.scores$person, team.scores$score)
  
  
  # df with stuff I want to format on
  df.format <- df.scores %>%
    select(pick = Participant, Total) %>%
    left_join(df.elim, by = "pick") %>%
    mutate(rank = rank(-Total, ties.method = "min"),
           font_face = ifelse(rank <= 3, "bold", "plain"),
           font_color = case_when(
             rank == 1 ~ "#D6AF36", # gold
             rank == 2 ~ "#A7A7AD", # silver
             rank == 3 ~ "#A77044", # bronze
             eliminated > 900 ~ "black",
             TRUE ~ "gray65"
           ),
           pick.temp = ifelse(rank <= 3, paste0("**", pick, "**"), pick),
           pick.label = glue("<span style='color:{font_color}'>{pick.temp}</span>")
           ) %>%
    select(-Total)
  
  df.format2 <- team.scores %>%
    mutate(font_face = ifelse(rank <= 3, "bold", "plain"),
           font_color = case_when(
             rank == 1 ~ "#D6AF36", # gold
             rank == 2 ~ "#A7A7AD", # silver
             rank == 3 ~ "#A77044", # bronze
             TRUE ~ "black"
           ),
           person.temp = ifelse(rank <= 3, paste0("**", person, "**"), person),
           person.label = glue("<span style='color:{font_color}'>{person.temp}</span>")
           ) %>%
    select(-rank) %>%
    rename(team.score = score)

  # create main df: 
    # long version of df2 (1 row per person/pick/week)
    # merge in elimination week data
  df2l <- df2 %>%
    pivot_longer(-c(person, pick, Total), names_to = "week", values_to = "score") %>%
    # left_join(df.elim) %>%
    mutate(pick = factor(pick, levels = levels(bguys.order)),
           person = factor(person, levels = team.order),
           wk_index = as.numeric(str_extract(week, "\\d"))) %>%
    drop_na(score) %>% # drop weeks that haven't happened yet (or person eliminated)
    left_join(df.format, by = "pick") %>%
    left_join(df.format2, by = "person")
  
  # update score df
  df.scores2 <- df.scores %>%
    left_join(df.format, by = c("Participant" = "pick"))


# define certain useful values like the max score, number of weeks so far
max.bguy.score <- max(df.scores$Total)
max.team.score <- max(team.scores$score)

week.list <- df2l %>%
  drop_na(score) %>%
  pull(week) %>%
  unique()

# colors to use for plotting
  # named vector so can have consistent color per week
  # I think there are more than 9 weeks so I'll have to update this
  # chose some color palette I found online that seemed OK for now
colors9 <- c("#fd7f6f", "#7eb0d5", "#b2e061", "#bd7ebe", "#ffb55a", "#ffee65", "#beb9db", "#fdcce5", "#8bd3c7")
names(colors9) <- paste("Week", 1:9)

# for the 17 teams
  # https://sashamaps.net/docs/resources/20-colors/
colors17 <- c('#e6194B', '#3cb44b', '#4363d8', '#f58231', '#911eb4', '#42d4f4', '#f032e6', '#ffe119', '#fabed4', '#469990', '#dcbeff', '#9A6324', '#800000', '#808000', '#ffd8b1', '#000075', '#a9a9a9')
names(colors17) <- unique(df2$person)

# plot themes 
theme_bar <- function(){
  theme_bw() +
    theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 13, margin = margin(4, 0, 4, 0)),
        axis.text.x = element_text(size = 13),
        # axis.text.y = element_(face = df.pick$font_face, colour = df.pick$font_color),
        axis.text.y = element_markdown(size = 12),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 13, margin = margin(10, 0, 0, 0)),
        # legend.position = c(.78, .5),
        plot.title = element_markdown(size = 16),
        plot.subtitle = element_markdown(size = 13),
        panel.border = element_blank())
}


```

```{r reactive}

# define reactive elements for shiny
  # page 1: the 2 isoalted team plots + how many weeks to show
  person1 <- reactive(input$person_select1)
  person2 <- reactive(input$person_select2)
  week_select <- reactive(input$week_select)
  
  # page 2: which team lines to show
  people_select <- reactive(input$people_select)

```


# Bar Charts

## Sidebar {.sidebar data-width=160}


<!-- #### **Person 1** -->
```{r personInput}

# sidebar is the UI elemennt that contains the inputs for reactive stuff

# dropdown for the top individual team plot
selectInput(inputId = "person_select1",
            label = "Person 1",
            choices = sort(as.character(team.order)),
            selected = team.order[1],
            selectize = FALSE,
            )

# dropdown for the bottom individual team plot
selectInput(inputId = "person_select2",
            label = "Person 2",
            choices = sort(as.character(team.order)),
            selected = team.order[2],
            selectize = FALSE,
            )

# selection box to pick any combination of weeks
selectInput(inputId = "week_select",
            label = "Weeks",
            choices = week.list,
            selected = week.list,
            multiple = TRUE,
            selectize = FALSE,
            )

```

<!-- some example descriptive text, here just describes how to select multiple -->
<!-- To select multiple weeks:<br> -->
<!-- *shift+click* or<br> -->
<!-- *cmd+click* (Mac) or<br> -->
<!-- *ctrl+click* (Windows)<br><br> -->
<!-- (default is all weeks) -->

**Point system:**  

**5** = group date invite  

**10** = one-on-one date invite  

**20** = rose during rose ceremony  

**25** = one-on-one date rose  

**30** =  group date rose  

**Various** one-time bonus points  


Column {data-width=500}
-----------------------------------------------------------------------

### 

```{r, fig.height = 5, fig.width = 10}

# reactive elements need to go within a "render" function
renderPlot({
  
  # get the ranking of the selected team 
    # reactive input is called as a function, e.g. person1()
    # will return the person (team) that is selected
  temp <- team.scores %>%
    filter(person == person1()) %>%
    pull(rank)

  # turn rank into a label to use in the plot text, e.g. 1st place, 4th place
  rank.label <- paste0("**", scales::ordinal(temp), "** place")

  # and "Tied for" if there is a tie for that ranking
  if(sum(team.scores$rank == temp) > 1){
    rank.label <- paste("Tied for", rank.label)
  }

  # get the team's score to use in the plot text
  score.label <- team.scores %>%
    filter(person == person1()) %>%
    pull(score)

  # bar plot of the points acquired by each of the team's pick
    # with fill color mapped to week, week is reactive
      # by default all weeks are selected (i.e. whole season so far)
      # but can select individual weeks or some other subset
    # horixontal bars so we can have the bguys names horizontal on y axis
    # no legend, so that we don't repeat the legend for each plot on the page,
      # because it is the same each time
    # title with the person's name
    # subtitle with their ranking and points
  df2l %>%
    filter(person == person1(),
           week %in% week_select()) %>%
  ggplot(aes(pick.label, score, fill = week)) +
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
    geom_hline(yintercept=0, size = 1.2, color = "black") + 
    coord_flip() +
    scale_y_continuous(limits = c(-.1, 1.1*max.bguy.score), expand = c(0,0)) +
    theme_bar() +
    theme(legend.position = "none") +
    scale_fill_manual(values = colors9) + 
    labs(x = NULL,
         y = NULL,
         title = paste0(person1(), "'s Team"),
         subtitle = paste0(rank.label, " with **", score.label, "** points"))

})



# # for testing stuff out, replaces reactive parts with static
#   temp <- team.scores %>%
#   filter(person == team.order[1]) %>%
#   pull(rank)
# 
#   rank.label <- paste(scales::ordinal(temp), "place")
# 
#   if(sum(team.scores$rank == temp) > 1){
#     rank.label <- paste("Tied for", rank.label)
#   }
# 
#   score.label <- team.scores %>%
#     filter(person == team.order[1]) %>%
#     pull(score)
# 
#   # # looking at different color palettes
#   #   # there is never anything I like
#   # colors.br <- RColorBrewer::brewer.pal("Set2", n = 12)
#   # 
#   # RColorBrewer::display.brewer.pal("Blues", n = 5)
#   # RColorBrewer::display.brewer.pal("Oranges", n = 6)
#   # RColorBrewer::display.brewer.all()
#   # 
#   # colors1 <- rev(RColorBrewer::brewer.pal("Blues", n = 5))
#   # colors2 <- rev(RColorBrewer::brewer.pal("Oranges", n = 6)[1:5])
#   # colors.BlO <- c(rbind(colors1, colors2))
# 
#   # some pastel paelette I found online
#   colors9 <- c("#fd7f6f", "#7eb0d5", "#b2e061", "#bd7ebe", "#ffb55a", "#ffee65", "#beb9db", "#fdcce5", "#8bd3c7")
#   names(colors9) <- paste("Week", 1:9)
# 
# df2l %>%
#   mutate(week = factor(week)) %>%
#   filter(person == team.order[1]) %>%
# ggplot(aes(pick.label, score, fill = week)) +
#   geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
#   coord_flip(ylim = c(0, max.bguy.score+5)) +
#   theme_bar() +
#   # theme(legend.title = element_blank(),
#   #       axis.text.y = element_markdown(size = 12)) +
#   # scale_fill_brewer(type = "qual", palette = 3) +
#   # scale_fill_manual(values = colors.BlO) +
#   scale_fill_manual(values = colors9) +
#   labs(x = NULL,
#        y = "Score",
#        title = paste0(team.order[1], "'s Team"),
#        subtitle = paste(rank.label, "with", score.label, "points"))


```


### 

```{r fig.height = 5, fig.width = 10}



# reactive elements need to go within a "render" function
# same as Team 1 plot, except uses person2() instead of person1()
# comments here may become out of date, see Team 1 version instead
# actually I should probably make a function instead of copy-pasting the plot code
renderPlot({
  
  # get the ranking of the selected team 
    # reactive input is called as a function, e.g. person1()
    # will return the person (team) that is selected
  temp <- team.scores %>%
    filter(person == person2()) %>%
    pull(rank)

  # turn rank into a label to use in the plot text, e.g. 1st place, 4th place
  rank.label <- paste0("**", scales::ordinal(temp), "** place")

  # and "Tied for" if there is a tie for that ranking
  if(sum(team.scores$rank == temp) > 1){
    rank.label <- paste("Tied for", rank.label)
  }

  # get the team's score to use in the plot text
  score.label <- team.scores %>%
    filter(person == person2()) %>%
    pull(score)

  # bar plot of the points acquired by each of the team's pick
    # with fill color mapped to week, week is reactive
      # by default all weeks are selected (i.e. whole season so far)
      # but can select individual weeks or some other subset
    # horixontal bars so we can have the bguys names horizontal on y axis
    # no legend, so that we don't repeat the legend for each plot on the page,
      # because it is the same each time
    # title with the person's name
    # subtitle with their ranking and points
  df2l %>%
    filter(person == person2(),
           week %in% week_select()) %>%
  ggplot(aes(pick.label, score, fill = week)) +
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
    geom_hline(yintercept=0, size = 1.2, color = "black") + 
    coord_flip() +
    scale_y_continuous(limits = c(-.1, 1.1*max.bguy.score), expand = c(0,0)) +
    theme_bar() +
    theme(legend.position = "none") +
    scale_fill_manual(values = colors9) + 
    labs(x = NULL,
         y = NULL,
         title = paste0(person2(), "'s Team"),
         subtitle = paste0(rank.label, " with **", score.label, "** points"))

})

```

### {data-height=80}
```{r}

# adds the legend for the whole first page
# located in the 1st column below the 2 team plots
# legend for fill:week map

# get the current week, so only show up to this week
max_wk <- max(df2l$wk_index)

# create mostly barebones plot with this legend
temp <- df2l %>%
ggplot(aes(fct_rev(person), score, fill = week)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 14),
        legend.position = "bottom") +
  scale_fill_manual(values = colors9[1:max_wk])

# extract just the legend with cowplot package
# there are probably other ways to do this, but I've just done it this way before
# when using cowplot to combine multiple plots with plot_grid()
l <- cowplot::get_legend(temp)

# need to wrap in plot_grid() and renderPlot to ouput correctly
renderPlot(cowplot::plot_grid(l))

# OLD: thought I needed to use fillCol at first, but I don't
# leaving here as an example in case something like this is needed in the future
# output$legend <- renderPlot({
#   cowplot::plot_grid(l)
# })
# 
# fillCol(plotOutput("legend", height = "100%"), height = 50)

```

Column {.tabset data-width=500}
-----------------------------------------------------------------------

### All Teams

```{r, fig.height=12, fig.width = 6}

# second column, on right-hand side of page
# has two tabs to select from (shows one plot at a time, one plot per tab)
# "All Teams" plot is the first plot (first tab)
# shows all the teams at once in one horizontal bar chart
# broken down by week, but not by bguys 
# sorted from top scorer to bottom scorer

renderPlot({

  df2l %>%
      filter(week %in% week_select()) %>%
    
  ggplot(aes(reorder(person.label, team.score), score, fill = week)) +
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
    geom_hline(yintercept=0, size = 1.2, color = "black") + 
    coord_flip() +
    scale_y_continuous(limits = c(-.1, 1.1*max.team.score), expand = c(0,0)) +
    theme_bar() +
    theme(legend.position = "none") +
    scale_fill_manual(values = colors9) +
    labs(x = NULL,
         y = "Score")
})


```

### All Bachelorette Contestants

```{r, fig.height=12, fig.width = 6}

# second tab shows all the bguys and their scores per week
# sorted from top scorer to bottom scorer

scoresl <- df.scores2 %>%
  pivot_longer(matches("Week"), names_to = "week", values_to = "score")

renderPlot({

scoresl %>%
    filter(week %in% week_select()) %>%
ggplot(aes(reorder(pick.label, Total), score, fill = week)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  geom_hline(yintercept=0, size = 1.2, color = "black") + 
  coord_flip() +
  scale_y_continuous(limits = c(-.1, 1.1*max.bguy.score), expand = c(0,0)) +
  theme_bar() +
  theme(legend.position = "none") +
  scale_fill_manual(values = colors9) + 
  labs(x = NULL,
     y = "Score")
})


```

# Line Chart
## Sidebar {.sidebar data-width=160}


<!-- #### **Person 1** -->
```{r}

# sidebar for second page
# a select-multiple box, choose which teams to show

selectInput(inputId = "people_select",
            label = "People",
            choices = sort(as.character(team.order)),
            selected = sort(as.character(team.order)),
            multiple = TRUE,
            selectize = FALSE,
            size = 17,
            )

```

To select multiple people:<br>
*shift+click* or<br>
*cmd+click* (Mac) or<br>
*ctrl+click* (Windows)<br><br>
(default is all people)

## {data-width="100%"}
###
```{r fig.height = 9, fig.width = 12}

# plot team scores over time (week)
# instead of legend, label the lines at the end of the last week

# # get cumulative score across weeks
  # includes clunky method to get people who have 0 players remaining
  # to have 0 for the remaining weeks and same total score
df.lineplot <-df2l %>%
  group_by(person, week) %>%
  summarise(score = sum(score)) %>%
  pivot_wider(names_from = week, values_from=score) %>%
  pivot_longer(-person, names_to = "week", values_to = "score") %>%
  mutate(score = replace_na(score, 0)) %>%
  mutate(wk_index = as.numeric(str_extract(week, "\\d"))) %>%
  group_by(person) %>%
  mutate(total_score = cumsum(score)) %>%
  select(person, week, wk_index, score, total_score)

# will do x-axis as continuous, but then re-label to include the word "Week" 
week.labels <- paste("Week", 1:max_wk)

renderPlot({

  # filter to selected teams
  dfplot <- df.lineplot %>%
    filter(person %in% people_select())

  # could add names at left-hand side too
  # dflab1 <- dfplot %>%
  #   ungroup() %>%
  #   filter(wk_index == min(wk_index))
  
  # add names at right-hand side instead of legend
  dflab2 <- dfplot %>%
    ungroup() %>%
    filter(wk_index == max(wk_index))

  # 
  # dfplot %>%
  #   ggplot(aes(wk_index, total_score, color = person)) +
  #   geom_line(aes(group = person), alpha = .8) +
  #   geom_point(alpha = .5) +
  #   # ggrepel::geom_text_repel(data = dflab,
  #   #                          aes(label = person),
  #   #                          nudge_x = .1) +
  #   ggrepel::geom_text_repel(data = dflab2,
  #                          aes(label = person),
  #                          nudge_x = .1) +
  #   scale_x_continuous(limits = c(.95, 2.2),
  #                      breaks = c(1:max_wk), labels = week.labels) +
  #   coord_cartesian(ylim = c(0, max.team.score+5)) +
  #   theme_bw() +
  #   theme(legend.position = "none",
  #         panel.grid.minor.x = element_blank()) +
  #   labs(y = "Score",
  #        x = "Week")
  
  dfplot %>%
    ggplot(aes(wk_index, total_score, color = person)) +
    geom_line(aes(group = person), alpha = .8, size = .7) +
    geom_point(alpha = .6, size = 3) +
    # ggrepel::geom_text_repel(data = dflab,
    #                          aes(label = person),
    #                          nudge_x = .1) +
    ggrepel::geom_text_repel(data = dflab2,
                           aes(label = person),
                           nudge_x = .1, size = 6) +
    scale_x_continuous(limits = c(.95, max_wk+.5),
                       breaks = c(1:max_wk), labels = week.labels) +
    coord_cartesian(ylim = c(0, max.team.score+5)) +
    scale_color_manual(values = colors17) +
    theme_bw() +
    theme(panel.grid.major.x = element_blank(),
          # panel.grid.major.y = element_line(color = "gray90"),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 17),
          axis.text.y = element_text(size = 16),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(color = "black", size = .3),
          plot.title = element_markdown(size = 20)
          ) +
    theme(legend.position = "none",
          panel.grid.minor.x = element_blank()) +
    labs(y = NULL,
         x = NULL,
         title = "Team scores over time") 

})



# # for testing:
#   # includes clunky method to get people who have 0 players remaining
#   # to have 0 for the remaining weeks and same total score
# dfplot <- df2l %>%
#   group_by(person, week) %>%
#   summarise(score = sum(score)) %>%
#   pivot_wider(names_from = week, values_from=score) %>%
#   pivot_longer(-person, names_to = "week", values_to = "score") %>%
#   mutate(score = replace_na(score, 0)) %>%
#   mutate(wk_index = as.numeric(str_extract(week, "\\d"))) %>%
#   group_by(person) %>%
#   mutate(total_score = cumsum(score)) %>%
#   select(person, week, wk_index, score, total_score)
# 
# 
# dflab1 <- dfplot %>%
#   ungroup() %>%
#   filter(wk_index == min(wk_index))
# dflab2 <- dfplot %>%
#   ungroup() %>%
#   filter(wk_index == max(wk_index))
# 
# 
# dfplot %>%
#   ggplot(aes(wk_index, total_score, color = person)) +
#   geom_line(aes(group = person), alpha = .8, size = .7) +
#   geom_point(alpha = .6, size = 3) +
#   # ggrepel::geom_text_repel(data = dflab,
#   #                          aes(label = person),
#   #                          nudge_x = .1) +
#   ggrepel::geom_text_repel(data = dflab2,
#                          aes(label = person),
#                          nudge_x = .1, size = 6) +
#   scale_x_continuous(limits = c(.95, max_wk+.5),
#                      breaks = c(1:max_wk), labels = paste0("Week", 1:max_wk)) +
#   coord_cartesian(ylim = c(0, max.team.score+5)) +
#   scale_color_manual(values = colors17) +
#   theme_bw() +
#   theme(panel.grid.major.x = element_blank(),
#         # panel.grid.major.y = element_line(color = "gray90"),
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 13),
#         axis.text.y = element_text(size = 14),
#         axis.ticks = element_blank(),
#         panel.border = element_blank(),
#         axis.line = element_line(color = "black", size = .3),
#         plot.title = element_markdown(size = 15)
#         ) +
#   theme(legend.position = "none",
#         panel.grid.minor.x = element_blank()) +
#   labs(y = NULL,
#        x = NULL,
#        title = "Team scores over time")


```


```{r}
ui <- fluidPage(
)

server <- function(input, output, session) {
}

shinyApp(ui, server)
```
